Gametic Selection, Meiotic Drive, Sex Ratio Bias, and Transitions Between sex-determination systems

### 2017-07-10 LaTeX/Turnover.tex
discussion and Results finessing

Animals and angiosperms exhibit extremely diverse sex-determination systems \citep[reviewed in][]{Bull:1983vi,Charlesworth:2010it,Beukeboom:2014vb,Bachtrog:2014bx}. 
Among species with genetic sex determination of diploid sexes, some taxa have heterogametic males (XY) and homogametic females (XX), including %non-monotreme? 
mammals and most dioecious plants \citep{Ming:2011iy}; whereas other taxa have homogametic males (ZZ) and heterogametic females (ZW), including Lepidoptera and birds. 
Within several taxa, the chromosome that harbours the master sex-determining region changes. 
For example, transitions of the master sex-determining gene between chromosomes or the evolution of new master sex-determining genes have occurred in Salmonids \citep{Li:2011fm,Yano:2012di}, Diptera \citep{Vicoso:2015hf}, and \textit{Oryzias} \citep{Myosho:2012fv}.
%Presentation at evolution found neo sex chromosome in birds, doesn't seem to be published or bioRxiv yet (consider pers com.), title: A previously unknown neo-sex chromosome mediates plumage divergence and speciation in hybridizing birds Jason Sardell; Elizabeth Cooper; J. Albert C. Uy. I wonder if this is actually a fusion event though?
In addition, many gonochoric clades with genetic sex determination exhibit transitions between male (XY) and female (ZW) heterogamety, including lizards \citep{Ezaz:2009tk}, eight of 26 teleost fish families \citep{Mank:2006bt}, true fruit flies \citep[Tephritids,][]{Vicoso:2015hf}, amphibians \citep{Hillis:1990gu}, the angiosperm genus \textit{Silene} \citep{Slancarova:2013dq}, and Coleoptera and Hemiptera \citep[][plate 2]{Beukeboom:2014vb}.
Indeed, in some cases, both male and female heterogametic sex-determination systems can be found in the same species, as exhibited by some cichlid species \citep{Ser:2010iq} and \textit{Rana rugosa} \citep{Ogata:2007jm}.
In addition, multiple transitions have occurred between genetic and environmental sex-determination systems, e.g., in reptiles and fishes \citep{Conover:1987in,Mank:2006bt,Pokorna:2009ui,Ezaz:2009tk,Pen:2010kk,Holleley:2015hc}.

@@

%\textcolor{blue}{
In Figure \ref{fig:regionplot_ 2a}A we show the region of parameter space within which both neo-W haplotypes invade ($\lambda_{mA}>1$ and $\lambda_{ma}>1$) when there is overdominance in females and no haploid selection \citep[corresponding to Figure 2a in][]{Otto2014}. 
Wherever both haplotypes have positive growth rates (gray region of Figure \ref{fig:regionplot_ 2a}), invasion by a neo-W is expected regardless of its linkage with the selected locus (i.e., even unlinked neo-W alleles can invade). 
In regions where only one haplotype can spread (white region of Figure \ref{fig:regionplot_ 2a}), a neo-W can invade as long as equation \eqref{eq:lambdasGen} is satisfied, which can require that the recombination rate, $R$, is small enough and yet still indicates that more loosely linked sex-determining regions can spread.
It is also possible for haploid selection to drive the invasion of a loosely linked neo-W. 
Take for instance selection directionally favouring $A$ in both diploid sexes and meiotic drive in males.
Figure \ref{fig:regionplot_ 2a}B then shows that ploidally-antagonisitic selection can allow both neo-W haplotypes to invade. 
%}

@@

Equation \eqref{eq:lambda_neoW} shows that, in contrast to the tight linkage results of the previous section, if there is no haploid selection ($t^\Hermaphrodite=\alpha^\Hermaphrodite_{\Delta}=0$), as considered by \citet{vanDoorn:2010hu} and weak selection, the spread of a neo-W is equivalent to the spread of a neo-Y ($\lambda_{W',XY}=\lambda_{Y',XY}$), such that heterogametic transitions (XY to ZW or ZW to XY) can also occur only if the neo-sex-determining region is more closely linked to a locus under selection ($R<r$). 
However, if there is any haploid selection, the additional term in equation \eqref{eq:lambda_neoW} can be positive, which can allow, for example, neo-W invasion ($\lambda_{W',XY}>1$) even when the neo-sex-determining region is less closely linked to the selected locus ($R>r$). 
These transitions are unusual because, when $R>r$, associations that selection has built up between alleles more favourable in one sex and alleles that determine sex will be weakened. 
Mean diploid fitness therefore decreases during heterogametic transitions that create looser sex-linkage (Figure \ref{fig:Combination_MeanFit}B,D).

@@

Previous research suggests that when the ancestral sex-determining locus is linked to a locus that experiences haploid selection (e.g., meiotic drive), a new, unlinked sex-determining locus invades in order to restore equal sex ratios \citep{Kozielska:2010vm}. 
Consider, for example, the case where the \textbf{A} locus is linked to the ancestral-SDR ($r<1/2$) and experiences meiotic drive in males only (e.g., during spermatogenesis but not during oogenesis, $\alpha^\male \neq 1/2$, $\alpha^\female=1/2$). 
Disregard gametic competition ($t^\female=t^\male=0$) such that zygotic sex ratios are only biased by meiotic drive in males. 
In this case, the zygotic sex ratio can be initially biased only if the ancestral sex-determining system is XY (Figure \ref{fig:Combination_Turnover}B). 
%If the ancxestral sex-determining system is ZW, the zygotic sex ratio will be 1:1 because diploid sex is determined by the proportion of Z-bearing versus W-bearing eggs and meiosis in females is fair (Figure \ref{fig:Combination_Turnover}D).
We might therefore expect a difference in the potential for XY to ZW and ZW to XY transitions. 
However, with weak selection we find that invasion by a neo-W into an XY system and invasion by a neo-Y into a ZW system occur under the same conditions. 
As it turns out, with weak selection, the strength of a male biased zygotic sex ratio favouring the invasion of a neo-W is equal to the strength of meiotic drive favouring the invasion of a driving neo-Y are equal (at least up to order $\epsilon^2$).
As selection becomes stronger (or linkage becomes tighter), this symmetry between sex-ratio selection and drive is lost, causing differences in the strength of selection favouring the two heterogametic transitions (compare red to black near -25cM and 25 cM in Figure \ref{fig:Combination_Centimorgans}).

@@

Under Fisherian sex-ratio selection, autosomal modifiers favour equal investment in male and female offspring, i.e., a 1:1 sex ratio \citep{Fisher:1930wy,Charnov:1982wg,West:2009we}. 
A novel environmental sex-determiner that causes half of its carriers to become female and half to become male ($k=1/2$) will be in males half of the time and in females half of the time (like an autosome).
In addition, these novel sex-determination alleles equalize the sex ratio and therefore one might expect them to be favoured by Fisherian sex-ratio selection when the resident sex ratio is biased.
However, assuming weak selection, we find that the growth rate of a rare, dominant offspring-controlled neo-ESD allele that produces males or females with equal probability ($k=1/2$) is
\begin{equation}
\lambda_{ESD',XY} =1+ \frac{1}{2}\frac{(\lambda_{Y',XY}-1) + (\lambda_{W',XY}-1)}{2} \Big|_{R=1/2} + O\left(\epsilon^3\right),
%V_A \frac{r}{1-2r}  \frac{s_{f} \left( s^\male - 3 s^\female\right) }{8 \left( s_{f} + s_{m} \right)^2} t_{m}^2.
\label{eq:lambda_ESD}
\end{equation}
\noindent
where we have indicated that $\lambda_{Y',XY}$ and $\lambda_{W',XY}$ are evaluated at $R=1/2$.
That is, recombination between the selected locus and the novel sex-determining locus, $R$, doesn't enter into the $k=1/2$ results.
This is because sex is essentially randomized each generation, %as if the novel sex-determiner were on an unlinked site that half the time acts as a neo-$W$ and half the time as a neo-$Y$.
preventing associations from building up between allele $A$ and sex.

Equation \eqref{eq:lambda_ESD} shows that invasion by a novel `perfect' ESD (equal sex ratio, $k=1/2$) mutation is the same for an ancestrally XY or ZW system (since $\lambda_{Y',XY} = \lambda_{W',ZW}$, $\lambda_{W',XY} = \lambda_{Y',ZW}$).
Thus, by the same argument as above (if drive only occurs in males then the sex ratio is only biased when the ancestral sex-determination system is XY), Fisherian sex-ratio selection alone does not explain the invasion of an offspring-controlled neo-ESD allele under weak selection. 
Rather, the neo-ESD gets half of the fitness of a feminizing mutation (neo-$W$) and half of the fitness of a masculinizing mutation (neo-$Y$), but only has an effect one half of the time (the other half of the time it produces the same sex as the ancestral system would have, to leading order). 
The net result can be that perfect ESD will not invade, even if current sex ratios are biased. 
For example, if there is haploid selection in males (either drive or pollen/sperm competition) but the conditions in table \ref{tab:specialcases} are not met, perfect ESD will not invade, even though it would equalize the zygotic sex ratio from an initially biased case (assuming $r<1/2$).

@@

Fisherian sex-ratio selection is sometimes considered in terms of balancing parental investment in male versus female offspring \citep{Charnov:1982wg}.
In addition, under environmental sex-determination, the proportion of males/females is sometimes controlled by the mother (e.g., the proportion of eggs laid in warm versus cold environments). 
We therefore also considered the invasion of a neo-sex-determining allele ($m$) in a model in which mothers that have at least one $m$ allele produce daughters with probability $k$. 
As with offspring-controlled ESD, for all $k\in\{0,1/2,1\}$, we find that invasion into an ancestral XY system is the same as invasion into an ancestrally ZW system (at least up to order $\epsilon^3$, assuming weak selection), implying that transitions between genetic sex-determination and maternally controlled environmental sex-determination are not driven by Fisherian sex-ratio selection alone.
\textcolor{red}{(Maternal ESD analysis still lacks meiotic drive -- Mathematica can't seem to deal with the added complexity.)}

@@

%MAIN RESULT
Two predominant theories explaining the remarkably high frequency of transitions between sex-determination systems are sexually-antagonistic selection and sex-ratio selection \citep[reviewed in][]{Blaser2012}.
The former predicts that neo-sex-determining alleles can invade when they arise in closer linkage with a sexually-antagonistic locus \citep{vanDoorn:2007eu,vanDoorn:2010hu}.
The latter predicts that neo-W alleles will invade an XY system when there is a male bias caused by haploid selection in males, and vice-versa, a neo-Y will invade a ZW system when there is a female bias caused by haploid selection in females \citep{Kozielska:2010vm,Ubeda:2015fx}.
Here we have shown that both predictions must be amended when recombination is weak relative to selection or selection happens in both diploid and haploid phases.

%TIGHT LINKAGE RESULTS
When the rate of recombination between the ancestral sex-determining locus and a locus under selection is small relative to the strength of selection (i.e., sex-linkage is tight, or selection is strong), heterogametic transitions (XY to ZW or ZW to XY) that reduce sex-linkage are possible, with or without haploid selection or sexually-antagonistic selection (Figure \ref{fig:regionplot_2a}).
The likelihoods of these transitions are driven by sex-ratio selection, direct selection on alleles linked to the neo-sex-determining allele, the ability of the neo-sex-determining allele to avoid selection in one sex, and the ability of the neo-sex-determining allele to bring alleles on the sex-specific chromosome in the ancestor into the other sex (given that the neo-sex determining allele is epistatically dominant to its predecessor).
This possibility that looser sex-linkage could evolve, even in the absence of haploid selection (Figure \ref{fig:regionplot_2a}A), was overlooked in \cite{vanDoorn:2010hu}, likely because they did not explicitly calculate the resident equilibria \citep[equation \ref{eq:tightequil};][]{Lloyd1977,Otto2014}.
Interestingly, there is substantial overlap between the parameter space that allows both neo-W-$A$ and neo-W-$a$ haplotypes to spread in an XY system and that which selects for increased recombination between X and Y chromosomes \citep[e.g., compare gray region of Figure \ref{fig:regionplot_2a}A with coloured regions of Figure 2(a) in][]{Otto2014}.
This makes sense, as when both neo-W haplotypes can spread the neo-W can invade despite reducing sex-linkage, i.e., the rate of recombination between the sex-determining allele and the selected locus increases.
\textcolor{red}{explain difference too? also explain why a neo-Y doesn't invade to reduce linkage despite selection for more recombination? mention something about haploid selection (Fig7B)?} \textcolor{blue}{maybe leave for now}

%HAPLOID SELECTION CAN ACT LIKE SEXUALLY-ANTAGONISTIC SELECTION
%It has previously been demonstrated that new sex-determining systems can evolve if there is genetic variation maintained by sexually-antagonistic selection \citep{vanDoorn:2007eu,vanDoorn:2010hu}. 
%In particular, 
Under weak selection (or loose sex-linkage), transitions to new sex-determining systems can occur when they arise more closely linked to a sexually-antagonistic locus \citep{vanDoorn:2007eu,vanDoorn:2010hu}.
Our results show that genetic variation at loci that experience haploid selection can generate selection in favour of new sex-determining systems in a similar way. 
New sex-determining alleles are again favoured if they are more closely linked to a locus under haploid selection.
%HAPLOID SELECTION ALSO ALLOWS LOOSER LINKAGE TO EVOLVE
However, with haploid selection, heterogametic transitions (XY to ZW or ZW to XY) can also occur when the new sex-determining region is less closely linked to the locus under selection. 
Neo-W (neo-Y) alleles invade when their fitness in females (males) is greater than the mean fitness of females (males) under the ancestral sex-determination system.
With sexually-antagonistic selection (between diploid sexes) only, linkage between a selected locus and the sex-determining region strengthens associations between male beneficial alleles and the male-determining allele (Y or Z) and between female beneficial alleles and the female-determining allele (X or W). 
Thus, the mean fitness of both males and females increases with closer linkage to the sex-determining region. 
Therefore, new sex-determining alleles only invade if they are more closely linked than the ancestral sex-determining region. 
However, if there is haploid selection on loci linked to an XY (ZW) sex-determining region, selection can maintain polymorphisms at which the mean fitness of females (males) is lower than it would be without sex-linkage. 
In these cases, unlinked neo-W (neo-Y) alleles can increase the fitness of the only sex they are found in, at a cost to the other sex, and invade despite lowering mean fitness (Figure \ref{fig:Combination_MeanFit}). 
\textcolor{red}{This is similar to mitochondria and male disease ...?}

%\textcolor{blue}{thought: maybe we could say somewhere that our results suggest more instability in sex-determination system than previously thought. Particularly, we show that variation around the ancestral sex-determining locus could favour new sex-determination systems under both haploid and diploid sex-specific selection, in contrast to previous results that rely on new sex-determining alleles arising in linkage with loci under sex-specific selection. This makes potential for cycling (which you mention in the next paragraph wrt sex ratio bias. Could be suitable here or in the conclusion?}

@@

%\textcolor{blue}{Maybe we could go along the angle of ``under the weak recombination assumptions, there is no asymmetry between XY to ZW and ZW to XY, indicating that sex ratio selection does not dominate. Asymmetry can be larger when linkage is tight (e.g., Fig 2C) but under most circumstances we do not predict asymmetry between XY to ZW and ZW to XY transitions despite the presence/absence of sex ratio selection. 
%Thus, haploid selection can favour heterogametic transitions both via sex ratio selection and via the fitness of alleles that are associated with the neo-sex-determining allele; these selection pressures are often of equal magnitude.''}

Sex ratio biases caused by gametic competition or meiotic drive have been shown to exert Fisherian sex-ratio selection on various autosomal \citep{Stalker:1961th,Smith:1975ft,Frank:1989vl,Hough:2013uo,Ubeda:2015fx, Otto:2015va} and sex-linked \citep{Ubeda:2015fx} modifiers.
We find that sex-ratio biases caused by haploid selection can also affect transitions between sex-determining systems (e.g., see $\zeta$ terms in Table \ref{tab:haplotype_growth}). 
For instance, when an allele that drives in males is linked to an XY locus it will often become associated with the Y and therefore produce a male bias ($\zeta<1/2$).
This male bias increases the potential for a neo-W to invade (as we then have $(2\zeta)^{-1}>1$ in Table \ref{tab:haplotype_growth}), which can equalize the sex-ratio \citep[for a related example see][]{Ubeda:2015fx}.
%However, the neo-W may not equalize the sex-ratio (e.g., if the driving allele also drives in females or affects competition among eggs)
However, this sex-ratio selection can be overwhelmed when the driving allele has additional selective effects (e.g., when it is detrimental for male diploids but beneficial for female diploids; Table \ref{tab:specialcases}), preventing the neo-W from invading. %and equalizing the sex ratio.
Indeed, these additional selective effects can even favour transitions between sex-determining systems that create new sex-ratio biases.
For example, in an ancestral ZW system, an allele that drives only in males can allow a linked neo-Y to invade, despite the fact it creates a male bias (Figure \ref{fig:Combination_Turnover}C).
%\textcolor{blue}{and figure 1, where the end of one panel is the same as the beginning of another, however, this figure may change).}
%What we would like to stress is that sex-ratio selection alone cannot predict when new sex-determining systems can evolve.
Furthermore, with weak selection, there is no asymmetry between XY to ZW and ZW to XY transitions, indicating that sex-ratio selection does not dominate (i.e., the sex-ratio bias created by haploid selection impacts the spread of a neo-W into an XY system the same way it impacts the spread of a neo-Y into a ZW system with a 1:1 sex ratio). 
An asymmetry can develop when sex-linkage is tight (e.g., Figure \ref{fig:Combination_Centimorgans} near -25cM and 25cM) but under most circumstances we do not predict asymmetry between XY to ZW and ZW to XY transitions despite the presence/absence of sex ratio selection. 
Thus, haploid selection can favour heterogametic transitions both via sex-ratio selection and via the fitness of alleles that are associated with the neo-sex-determining allele, and these selection pressures are often of equal magnitude. 
%\textcolor{blue}{nice, pretty subtle argument but I think that's a good job.}

>>>>

Animals and angiosperms exhibit extremely diverse sex-determination systems \citep[reviewed in][]{Bull:1983vi,Charlesworth:2010it,Beukeboom:2014vb,Bachtrog:2014bx}. 
Among species with genetic sex determination of diploid sexes, some taxa have heterogametic males (XY) and homogametic females (XX), including %non-monotreme? 
mammals and most dioecious plants \citep{Ming:2011iy}; whereas other taxa have homogametic males (ZZ) and heterogametic females (ZW), including Lepidoptera and birds. 
Within several taxa, the chromosome that harbours the master sex-determining region changes. 
For example, transitions of the master sex-determining gene between chromosomes or the evolution of new master sex-determining genes have occurred in Salmonids \citep{Li:2011fm,Yano:2012di}, Diptera \citep{Vicoso:2015hf}, and \textit{Oryzias} \citep{Myosho:2012fv}.
%Presentation at evolution found neo sex chromosome in birds, doesn't seem to be published or bioRxiv yet (consider pers com.), title: A previously unknown neo-sex chromosome mediates plumage divergence and speciation in hybridizing birds Jason Sardell; Elizabeth Cooper; J. Albert C. Uy. I wonder if this is actually a fusion event though?
In addition, many gonochoric clades with genetic sex determination exhibit transitions between male (XY) and female (ZW) heterogamety, including snakes \textcolor{red}{(Gamble et al. 2017, Current Biology)}, lizards \citep{Ezaz:2009tk}, eight of 26 teleost fish families \citep{Mank:2006bt}, true fruit flies \citep[Tephritids,][]{Vicoso:2015hf}, amphibians \citep{Hillis:1990gu}, the angiosperm genus \textit{Silene} \citep{Slancarova:2013dq}, and Coleoptera and Hemiptera \citep[][plate 2]{Beukeboom:2014vb}.
Indeed, in some cases, both male and female heterogametic sex-determination systems can be found in the same species, as exhibited by some cichlid species \citep{Ser:2010iq} and \textit{Rana rugosa} \citep{Ogata:2007jm}.
In addition, multiple transitions have occurred between genetic and environmental sex-determination systems, e.g., in reptiles and fishes \citep{Conover:1987in,Mank:2006bt,Pokorna:2009ui,Ezaz:2009tk,Pen:2010kk,Holleley:2015hc}.

@@

%\textcolor{blue}{
In Figure \ref{fig:regionplot_ 2a}A we show the region of parameter space within which both neo-W haplotypes invade ($\lambda_{mA}>1$ and $\lambda_{ma}>1$) when there is overdominance in females and no haploid selection \citep[corresponding to Figure 2a in][]{Otto2014}. 
Wherever both haplotypes have positive growth rates (gray region of Figure \ref{fig:regionplot_ 2a}), invasion by a neo-W is expected regardless of its linkage with the selected locus (i.e., even unlinked neo-W alleles can invade). 
In regions where only one haplotype can spread (white region of Figure \ref{fig:regionplot_ 2a}), a neo-W can invade as long as equation \eqref{eq:lambdasGen} is satisfied, which can require that the recombination rate, $R$, is small enough and yet still indicates that more loosely linked sex-determining regions can spread.
It is also possible for haploid selection to drive the invasion of a loosely linked neo-W. 
Take, for instance, selection directionally favouring $A$ in both diploid sexes and meiotic drive in males.
Figure \ref{fig:regionplot_ 2a}B then shows that ploidally-antagonisitic selection can allow both neo-W haplotypes to invade. 
%}

@@

Equation \eqref{eq:lambda_neoW} shows that, in contrast to the tight linkage results of the previous section, with weak selection and no haploid selection ($t^\Hermaphrodite=\alpha^\Hermaphrodite_{\Delta}=0$), as considered by \citet{vanDoorn:2010hu}, the spread of a neo-W is equivalent to the spread of a neo-Y ($\lambda_{W',XY}=\lambda_{Y',XY}$), such that heterogametic transitions (XY to ZW or ZW to XY) can also occur only if the neo-sex-determining region is more closely linked to a locus under selection ($R<r$). 
However, if there is any haploid selection, the additional term in equation \eqref{eq:lambda_neoW} can be positive, which can allow, for example, neo-W invasion ($\lambda_{W',XY}>1$) even when the neo-sex-determining region is less closely linked to the selected locus ($R>r$). 
These transitions are unusual because, when $R>r$, associations that selection has built up between alleles more favourable in one sex and alleles that determine sex will be weakened. 
Mean diploid fitness therefore decreases during heterogametic transitions that create looser sex-linkage (Figure \ref{fig:Combination_MeanFit}B,D).

@@

Previous research suggests that when the ancestral sex-determining locus is linked to a locus that experiences haploid selection (e.g., meiotic drive), a new, unlinked sex-determining locus invades in order to restore equal sex ratios \citep{Kozielska:2010vm}. 
Consider, for example, the case where the \textbf{A} locus is linked to the ancestral-SDR ($r<1/2$) and experiences meiotic drive in males only (e.g., during spermatogenesis but not during oogenesis, $\alpha^\male \neq 1/2$, $\alpha^\female=1/2$). 
Disregarding gametic competition ($t^\female=t^\male=0$) such that zygotic sex ratios are only biased by meiotic drive in males. 
In this case, the zygotic sex ratio can be initially biased only if the ancestral sex-determining system is XY (Figure \ref{fig:Combination_Turnover}B). 
%If the ancxestral sex-determining system is ZW, the zygotic sex ratio will be 1:1 because diploid sex is determined by the proportion of Z-bearing versus W-bearing eggs and meiosis in females is fair (Figure \ref{fig:Combination_Turnover}D).
We might therefore expect a difference in the potential for XY to ZW and ZW to XY transitions. 
However, to leading order with selection weak relative to recombination, we find that sex ratio selection (first terms in table \ref{tab:haplotype_growth}) is equal in magnitude to the fitness effects of alleles associated with new sex-determining alleles (second terms in table \ref{tab:haplotype_growth}).
Thus, invasion by a neo-W into an XY system and invasion by a neo-Y into a ZW system occur under the same conditions ($\lambda_{Y',XY}=\lambda_{W',ZW}$ and $\lambda_{Y',ZW}=\lambda_{W',XY}$, at least up to order $\epsilon^2$).
As selection becomes stronger (or linkage becomes tighter), this symmetry between sex-ratio selection and haploid selection is lost, causing differences in the strength of selection favouring the two heterogametic transitions (compare red to black near -25cM and 25 cM in Figure \ref{fig:Combination_Centimorgans}).

@@

%Under Fisherian sex-ratio selection, autosomal modifiers favour equal investment in male and female offspring, i.e., a 1:1 sex ratio \citep{Fisher:1930wy,Charnov:1982wg,West:2009we}. 
%A novel environmental sex-determiner that causes half of its carriers to become female and half to become male ($k=1/2$) will be in males half of the time and in females half of the time (like an autosome).
%In addition, these novel sex-determination alleles equalize the sex ratio and therefore one might expect them to be favoured by Fisherian sex-ratio selection when the resident sex ratio is biased.
%However, assuming weak selection, we find that the growth rate of a rare, dominant offspring-controlled neo-ESD allele that produces males or females with equal probability ($k=1/2$) is
Of particular interest are ESD mutations that cause half of their carriers to develop as females and half as males ($k=1/2$, creating equal sex ratios), the spread of which is given by
\begin{equation}
\lambda_{ESD',XY} =1+ \frac{1}{2}\frac{(\lambda_{Y',XY}-1) + (\lambda_{W',XY}-1)}{2} \Big|_{R=1/2} + O\left(\epsilon^3\right),
%V_A \frac{r}{1-2r}  \frac{s_{f} \left( s^\male - 3 s^\female\right) }{8 \left( s_{f} + s_{m} \right)^2} t_{m}^2.
\label{eq:lambda_ESD}
\end{equation}
\noindent
where we have indicated that $\lambda_{Y',XY}$ and $\lambda_{W',XY}$ are evaluated at $R=1/2$.
That is, recombination between the selected locus and the novel sex-determining locus, $R$, doesn't enter into the $k=1/2$ results.
This is because sex is essentially randomized each generation, %as if the novel sex-determiner were on an unlinked site that half the time acts as a neo-$W$ and half the time as a neo-$Y$.
preventing associations from building up between allele $A$ and sex.  
An important result from equation \eqref{eq:lambda_ESD} is that ESD can invade if there is haploid selection. 
When evaluated at $R=1/2$, $\lambda_{Y',XY} \leq 1$ but $\lambda_{W',XY}$ can be greater than one if there is haploid selection, as discussed above. 
Previous studies where ESD is favoured have typically assumed that environmental conditions (e.g., maternal condition, mate quality, age, or host size) can differentially affect the fitness of males versus females such that ESD invades because it allows sex determination to depend on the environment \citep[reviewed in][]{Charnov:1982wg,Bull:1983vi,West:2009we}. 
Here, ESD mutations can spread because they generate females that are either rare or have high fitness, in the same manner as a neo-W.

Equation \eqref{eq:lambda_ESD} also shows that invasion by a novel `perfect' ESD (equal sex ratio, $k=1/2$) mutation is the same for an ancestrally XY or ZW system (since $\lambda_{Y',XY} = \lambda_{W',ZW}$, $\lambda_{W',XY} = \lambda_{Y',ZW}$).
Thus, by the same argument as above (if drive only occurs in males then the sex ratio is only biased when the ancestral sex-determination system is XY), Fisherian sex-ratio selection alone does not explain the invasion of an offspring-controlled neo-ESD allele under weak selection. 
Rather, the neo-ESD gets half of the fitness of a feminizing mutation (neo-$W$) and half of the fitness of a masculinizing mutation (neo-$Y$), but only has an effect one half of the time (the other half of the time it produces the same sex as the ancestral system would have, to leading order). 
The net result can be that perfect ESD will not invade, even if current sex ratios are biased. 
For example, if there is haploid selection in males (either drive or pollen/sperm competition) but the conditions in table \ref{tab:specialcases} are not met, perfect ESD will not invade, even though it would equalize the zygotic sex ratio from an initially biased case (assuming $r<1/2$).

@@

%Fisherian sex-ratio selection is sometimes considered in terms of balancing parental investment in male versus female offspring \citep{Charnov:1982wg}.
%In addition, under environmental sex-determination, the proportion of males/females is sometimes controlled by the mother (e.g., the proportion of eggs laid in warm versus cold environments). 
%We therefore also considered the invasion of a neo-sex-determining allele ($m$) in a model in which mothers that have at least one $m$ allele produce daughters with probability $k$. 
%As with offspring-controlled ESD, for all $k\in\{0,1/2,1\}$, we find that invasion into an ancestral XY system is the same as invasion into an ancestrally ZW system (at least up to order $\epsilon^3$, assuming weak selection), implying that transitions between genetic sex-determination and maternally controlled environmental sex-determination are not driven by Fisherian sex-ratio selection alone.
%\textcolor{red}{(Maternal ESD analysis still lacks meiotic drive -- Mathematica can't seem to deal with the added complexity.)}

@@

%MAIN RESULT
Two predominant theories explaining the remarkably high frequency of transitions between sex-determination systems are sexually-antagonistic selection and sex-ratio selection \citep[reviewed in][]{Blaser2012} \textcolor{red}{(van Doorn, 2014, sexual development)}.
The former predicts that neo-sex-determining alleles can invade when they arise in closer linkage with a sexually-antagonistic locus \citep{vanDoorn:2007eu,vanDoorn:2010hu}.
The latter predicts that neo-W alleles will invade an XY system when there is a male bias caused by haploid selection in males, and vice-versa, a neo-Y will invade a ZW system when there is a female bias caused by haploid selection in females \citep{Kozielska:2010vm,Ubeda:2015fx}.
Here we have shown that both predictions must be amended when recombination is weak relative to selection or selection happens in both diploid and haploid phases.

%TIGHT LINKAGE RESULTS
When the rate of recombination between the ancestral sex-determining locus and a locus under selection is small relative to the strength of selection (i.e., sex-linkage is tight, or selection is strong), heterogametic transitions (XY to ZW or ZW to XY) that reduce sex-linkage are possible, with or without haploid selection or sexually-antagonistic selection (Figure \ref{fig:regionplot_2a}).
The likelihoods of these transitions are driven by sex-ratio selection, direct selection on alleles linked to the neo-sex-determining allele, the ability of the neo-sex-determining allele to avoid selection in one sex, and the ability of the neo-sex-determining allele to bring alleles on the sex-specific chromosome in the ancestor into the other sex (given that the neo-sex determining allele is epistatically dominant to its predecessor).
This possibility that looser sex-linkage could evolve, even in the absence of haploid selection (Figure \ref{fig:regionplot_2a}A), was overlooked in \cite{vanDoorn:2010hu}, likely because they did not explicitly calculate the resident equilibria \citep[equation \ref{eq:tightequil};][]{Lloyd1977,Otto2014}.
Interestingly, there is substantial overlap between the parameter space that allows both neo-W-$A$ and neo-W-$a$ haplotypes to spread in an XY system and that which selects for increased recombination between X and Y chromosomes \citep[e.g., compare gray region of Figure \ref{fig:regionplot_2a}A with coloured regions of Figure 2(a) in][]{Otto2014}.
This makes sense, as when both neo-W haplotypes can spread the neo-W can invade despite reducing sex-linkage, i.e., the rate of recombination between the sex-determining allele and the selected locus increases.
%\textcolor{red}{explain difference too? also explain why a neo-Y doesn't invade to reduce linkage despite selection for more recombination? mention something about haploid selection (Fig7B)?} \textcolor{blue}{maybe leave for now}

%HAPLOID SELECTION CAN ACT LIKE SEXUALLY-ANTAGONISTIC SELECTION
%It has previously been demonstrated that new sex-determining systems can evolve if there is genetic variation maintained by sexually-antagonistic selection \citep{vanDoorn:2007eu,vanDoorn:2010hu}. 
%In particular, 
Under weak selection (or loose sex-linkage), transitions to new sex-determining systems can occur when they arise more closely linked to a sexually-antagonistic locus \citep{vanDoorn:2007eu,vanDoorn:2010hu}.
Our results show that genetic variation at loci that experience haploid selection can generate selection in favour of new sex-determining systems in a similar way. 
New sex-determining alleles are again favoured if they are more closely linked to a locus under haploid selection.
%HAPLOID SELECTION ALSO ALLOWS LOOSER LINKAGE TO EVOLVE
However, with haploid selection, heterogametic transitions (XY to ZW or ZW to XY) can also occur when the new sex-determining region is less closely linked to the locus under selection. 
Neo-W (neo-Y) alleles invade when their fitness in females (males) is greater than the mean fitness of females (males) under the ancestral sex-determination system and/or females (males) are the rarer sex.
With sexually-antagonistic selection (between diploid sexes) only, linkage between a selected locus and the sex-determining region strengthens associations between male beneficial alleles and the male-determining allele (Y or Z) and between female beneficial alleles and the female-determining allele (X or W). 
Thus, the mean fitness of both males and females increases with closer linkage to the sex-determining region. 
Therefore, new sex-determining alleles only invade if they are more closely linked than the ancestral sex-determining region. 
However, if there is haploid selection on loci linked to an XY (ZW) sex-determining region, selection can maintain polymorphisms at which the product of the frequency of females (males) and the mean fitness of females (males) is lower than it would be without sex-linkage. 
In these cases, unlinked neo-W (neo-Y) alleles can increase the frequency and/or fitness of the only sex they are found in, at a cost to the other sex, and invade despite lowering population mean fitness (Figure \ref{fig:Combination_MeanFit}). 
%\textcolor{red}{This is similar to mitochondria and male disease ...?}

@@

Sex ratio biases caused by gametic competition or meiotic drive have been shown to exert Fisherian sex-ratio selection on various autosomal \citep{Stalker:1961th,Smith:1975ft,Frank:1989vl,Hough:2013uo,Ubeda:2015fx, Otto:2015va} and sex-linked \citep{Ubeda:2015fx} modifiers.
We find that sex-ratio biases caused by haploid selection can also affect transitions between sex-determining systems (e.g., see $\zeta$ terms in Table \ref{tab:haplotype_growth}). 
For instance, when an allele that drives in males is linked to an XY locus it will often become associated with the Y and therefore produce a male bias ($\zeta<1/2$).
This male bias increases the potential for a neo-W to invade (as we then have $(2\zeta)^{-1}>1$ in Table \ref{tab:haplotype_growth}), which can equalize the sex-ratio \citep[for a related example see][]{Ubeda:2015fx}.
%However, the neo-W may not equalize the sex-ratio (e.g., if the driving allele also drives in females or affects competition among eggs)
However, this sex-ratio selection can be overwhelmed when the driving allele has additional selective effects (e.g., when it is beneficial for male diploids but detrimental for female diploids; Table \ref{tab:specialcases}), preventing the neo-W from invading. %and equalizing the sex ratio.
Indeed, these additional selective effects can even favour transitions between sex-determining systems that create new sex-ratio biases.
For example, in an ancestral ZW system, an allele that drives only in males can allow a linked neo-Y to invade, despite the fact it creates a male bias (Figure \ref{fig:Combination_Turnover}C).
%\textcolor{blue}{and figure 1, where the end of one panel is the same as the beginning of another, however, this figure may change).}
%What we would like to stress is that sex-ratio selection alone cannot predict when new sex-determining systems can evolve.
Furthermore, with weak selection, there is no asymmetry between XY to ZW and ZW to XY transitions, indicating that sex-ratio selection does not dominate (i.e., the sex-ratio bias created by haploid selection impacts the spread of a neo-W into an XY system the same way it impacts the spread of a neo-Y into a ZW system with a 1:1 sex ratio). 
An asymmetry can develop when sex-linkage is tight (e.g., Figure \ref{fig:Combination_Centimorgans} near -25cM and 25cM) but under most circumstances we do not predict asymmetry between XY to ZW and ZW to XY transitions despite the presence/absence of sex ratio selection. 
Thus, haploid selection can favour heterogametic transitions both via sex-ratio selection and via fitness effects of alleles that are associated with the neo-sex-determining allele, and these selection pressures are often of equal magnitude. 
%\textcolor{blue}{nice, pretty subtle argument but I think that's a good job.}
